---
title: "Exercices in linear models and experimental design"
author: "Aina Montalbán"
output:
  html_document:
    theme: united
    toc: yes
  pdf_document:
    toc: no
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction

These exercises describe experimental situations that have been studied using some omics technology like microarrays.

# Comparison between groups

This case study is based on a data are available in GEO as series GSE124927 series on the following link
https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi

The researchers studied RNA isolated from FACS-sorted uterine natural killer cells from Gd10.5 pregnancies in 10 control, 6 diet-induced obese and 4 diet-induced obese resistant mice. Therefore, there are three groups of mice:

- 10 control diet
- 6 diet-induced obsese
- 4 diet-induced obese resistant

Obviously this is an observational study but its analysis can be done using a linear model approach as well.

1. Identify the experimental factors and their levels.
2. Write down the design matrix associated with this study design.
3. Build the contrast matrix needed to compare each group type with the oher two, that are:

    1. "Control diet" vs “Diet-induced obese”
    2. "Control diet" vs “Diet-induced obese resistant”
    3. "Diet-induced obese" vs “Diet-induced obese resistant”
    

```{r readTargets}
require (limma)
targets <- read.csv("targets.txt", sep = "")

```



# Design Matrix
We can compute the design matrix in two ways: manually or using the limma package.
```{r designMatrix}
design<-matrix(
  c(1,0,1,0,0,0,0,0,1,1,1,1,0,0,1,0,0,1,1,1,
    0,1,0,1,1,0,0,1,1,0,0,0,1,1,0,0,0,0,0,0,
    0,0,0,0,1,1,0,0,0,0,0,0,0,0,0,1,1,0,0,0),
  nrow=20,byrow=F)
design2 <-model.matrix(~ 0+targets$shortName)
colnames(design)<-c("CD", "HF", "HF_RES")
rownames(design)<- targets$sampleName 
colnames(design2)<-c("CD", "HF", "HF_RES")
rownames(design2)<- targets$sampleName 
print(design2)
```
  

# Contrast Matrix

```{r contrastsMatrix}
cont.matrix1 <- makeContrasts( 
        CD.vs.HF = CD- HF, 
        HF.vs.HF_RES = HF-HF_RES, 
        CD.vs.HF_RES = CD-HF_RES,
        levels = design2)
cont.matrix1
```

